---
title: "SA 2"
author: "Cuerdo, Naomi Hannah A., Morillo, Jade Marco"
date: "2024-05-11"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
library(DataCombine)
library(anytime)
library(plyr)
library(dplyr)
library(dgof)
library(fitdistrplus)
library(VGAM)
library(data.table)
library(EnvStats)
library(ggplot2)
library(tsallisqexp)
library(poweRlaw)
library(fitur)
library(fitdistrplus)
```


# "Bitcoin Change Distribution Analysis"
#### Find out which probability distribution function best fits Bitcoin’s returns for trading data every minute, from January 1, 2012 to April 15, 2024, for Bitcoin quoted in United States dollars or the BTC/USD pair.

The data for 1 minute returns is accessible, so the data presented is for 1 day returns. 

##### Setting up the data frame
```{r read}
df_btc <- read.csv("C:/Users/naomi/Documents/SA2_Cuerdo_Morillo/1day_BTC_USD.csv")
summary(df_btc)
```

##### Histogram of Returns
```{r histogram}
df_btc$Change <- as.numeric(gsub("%", "", df_btc$Change)) / 100
x_limits <- c(-0.3, 0.3)
# Histogram
hist(df_btc$Change, 
     main = "Histogram of Percent Change", 
     xlab = "returns", 
     ylab = "frequency ", 
     col = alpha("purple", 0.5),  # Adjust transparency
     breaks = 500,  # Reduce breaks for better visibility
     xlim = x_limits)

```

Given the histogram, we can now proceed to conduct the Kolmogrov-Smirnov (KS) Test. The smallest KS resuly will be the best fit for the data. 

##### Kolmogrov-Smirnov Test

#####  Normal Distribution
```{r normal}
# Normal Distribution
cleaned_data <- na.omit(df_btc$Change)
fit <- fitdist(cleaned_data, "norm")
random_sample <- rnorm(length(cleaned_data), mean = fit$estimate[1], sd = fit$estimate[2])
cleaned_data <- jitter(cleaned_data)
ks_result1 <- ks.test(cleaned_data, random_sample)
print(ks_result1)
```

#####  T Distribution
```{r  t}
# T Distribution
random_sample <- rt(length(cleaned_data), length(cleaned_data)-1)
ks_result2 <- ks.test(cleaned_data, random_sample)
print(ks_result2)
```


#####  Laplace Distribution
```{r laplace}
lp_test <-rlaplace(length(na.omit(df_btc$Change)), 
                   location = mean(na.omit(df_btc$Change)), 
                   scale = sd(na.omit(df_btc$Change)))
jittered_data <- jitter(na.omit(df_btc$Change))
ks_result3 <- ks.test(jittered_data, lp_test)
print(ks_result3)

```

##### Tsallis Distribution
```{r}
df_test <- rtsal(length(na.omit(df_btc$Change)), mean(na.omit(df_btc$Change)), sd(na.omit(df_btc$Change)))
jittered_data <- jitter(na.omit(df_btc$Change))
ks_result4 <- ks.test(jittered_data, df_test)
print(ks_result4)

```

## Results
```{r results}
cat("Normal:", ks_result1$statistic, "\n")
cat("T- Distribution:", ks_result2$statistic, "\n")
cat("Laplace:", ks_result3$statistic, "\n")
cat("TSALLIS:", ks_result4$statistic, "\n")
```

Given that the Laplace has the smallest value, Laplace will be the best probability distribution to describe Bitcoin's returns for trading on a daily basis, followed by the Normal Distribution. 

For better visualiztion, here is the Graph for the Probability Distribution that describes the data:

```{r}

# Histogram
hist(df_btc$Change, 
     main = "Histogram of Bitcoin Returns", 
     xlab = "Returns", 
     ylab = "frequency", 
     col = alpha("purple", 0.5),
     breaks = 500,
     xlim = x_limits)

#Laplace
laplace_pdf <- function(x, mu, b) {
  exp(-abs(x - mu) / b) / (2 * b)
}
laplace_shift <- function(x) {
  laplace_pdf(x, mu = 0, b = 0.04) * length(df_btc$Change) * diff(range(df_btc$Change)) / 250
}

curve(laplace_shift(x),
      from = x_limits[1], 
      to = x_limits[2], 
      add = TRUE, 
      col = "darkgreen", 
      lwd = 2)
#Normal 
shifted_normal_pdf <- function(x) {
  dnorm((x) / 0.06) * length(df_btc$Change) * diff(range(df_btc$Change)) / 10
}
curve(shifted_normal_pdf(x), 
      from = x_limits[1], 
      to = x_limits[2], 
      add = TRUE, 
      col = "blue", 
      lwd = 2)
# T Distribution
curve(dt(x, length(cleaned_data) - 1), 
      from = x_limits[1], 
      to = x_limits[2], 
      add = TRUE, 
      col = "red", 
      lwd = 2)
#Tsallis Distribution
curve(dtsal(x, mean(na.omit(df_btc$Change)), sd(na.omit(df_btc$Change))), 
      from = x_limits[1], 
      to = x_limits[2], 
      add = TRUE, 
      col = "orange", 
      lwd = 2)

legend("topright",
       legend = c("Histogram", "Laplace", "Normal", "T Distribution", "Tsallis Distribution"),
       col = c((alpha("purple", 0.5)), "darkgreen", "blue", "red", "orange"),
       lwd = c(10, 2, 2, 2, 2), 
       lty = c(1, 1, 1, 1, 1), 
       cex = 0.8)
```

From here, it is evident that Laplace best fits the data of Bitcoin Returns on a daily basis. 